Cross-Loop Optimization of Arithmetic Intensity for Finite Element Local Assembly

### 2014-09-03 TACO-paper-v2/coffee.tex
Adjust conclusions

In this paper, we have presented the design, optimizations and systematic
performance evaluation of COFFEE, a compiler for finite element local
assembly. In this context, to the best of our knowledge, COFFEE is the
first compiler capable of introducing low-level optimizations to
maximize instruction-level parallelism, register locality and SIMD
vectorization. Assembly kernels have particular characteristics. Their
iteration space is usually very small, with the size depending on
aspects like the degree of accuracy one wants to reach (polynomial
order of the method) and the mesh discretization employed. The data
space, in terms of number of arrays and scalars required to evaluate
the element matrix, grows proportionally with the complexity of the
finite element problem. COFFEE has been developed taking into account
all of these degrees of freedom, based on the idea that reducing the
problem of local assembly optimization to a fixed sequence of
transformations is far too superficial if close-to-peak performance
needs to be reached. The various optimizations overcome limitations of
current vendor and research compilers. The exploitation of domain
knowledge allows some of them to be particularly effective, as
demonstrated by our experiments on two state-of-the-art Intel
platforms. Further work includes a comprehensive study about
feasibility and constraints on transforming kernels into a sequence of
calls to external linear algebra libraries. COFFEE supports all of the
problems expressible in Firedrake, and is already integrated with
this framework. The generality and the applicability of the proposed 
code transformations to other domains has also been discussed.

>>>>

In this paper, we have presented the study and systematic performance 
evaluation of a class of composable cross-loop optimizations for improving 
arithmetic intensity in finite element local assembly kernels, and their 
integration in a novel compiler, COFFEE. In the context of automated code 
generation for finite element local assembly, COFFEE is the first compiler 
capable of introducing low-level optimizations to maximize instruction-level 
parallelism, register locality and SIMD vectorization. Assembly kernels have 
particular characteristics. Their iteration space is usually very small, with 
the size depending on aspects like the degree of accuracy one wants to 
reach (polynomial order of the method) and the mesh discretization employed. 
The data space, in terms of number of arrays and scalars required to evaluate
the element matrix, grows proportionally with the complexity of the
finite element problem. COFFEE has been developed taking into account
all of these degrees of freedom, based on the idea that reducing the
problem of local assembly optimization to a fixed sequence of
transformations is far too superficial if close-to-peak performance
needs to be reached. The various optimizations overcome limitations of
current vendor and research compilers. The exploitation of domain
knowledge allows some of them to be particularly effective, as
demonstrated by our experiments on two state-of-the-art Intel
platforms. Further work includes a comprehensive study about
feasibility and constraints on transforming kernels into a sequence of
calls to external linear algebra libraries. COFFEE supports all of the
problems expressible in Firedrake, and is already integrated with
this framework. The generality and the applicability of the proposed 
code transformations to other domains has also been discussed.
